---
title: "How COVID-19 pandemic changed emotions"
description: |
site: distill::distill_website
---

### A survey measured people's feelings, on 7 different emotions, during the COVID-19 pandemic in the U.S. How did people's feelings change during different periods of time?

## Details to help you to understand the graphs
+ Week 1: mid-November, 2019 
+ Purple line 1: Week 20 -- The beginning of the lockdown
+ Purple line 2: Week 30 -- The end of the lockdown
+ Red line: Peak of COVID deaths

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggthemes)
library(ggplot2)
library(tidyverse)
library(dplyr)
emotion_data_v2 <- read_csv("emotion data v2.csv")
```

## Sadness
```{r}
emotion_data_v2$emotions_yesterday_1[emotion_data_v2$emotions_yesterday_1 == 2]<-0 
   emotion_data_v2$emotions_yesterday_2[emotion_data_v2$emotions_yesterday_2 == 2]<-0
   emotion_data_v2$emotions_yesterday_3[emotion_data_v2$emotions_yesterday_3 == 2]<-0
   emotion_data_v2$emotions_yesterday_4[emotion_data_v2$emotions_yesterday_4 == 2]<-0
   emotion_data_v2$emotions_yesterday_5[emotion_data_v2$emotions_yesterday_5 == 2]<-0
   emotion_data_v2$emotions_yesterday_6[emotion_data_v2$emotions_yesterday_6 == 2]<-0
   emotion_data_v2$emotions_yesterday_7[emotion_data_v2$emotions_yesterday_7 == 2]<-0
   emotion_data_v2$older<-"55 or older"
   emotion_data_v2$older[emotion_data_v2$age<55]<-"30- 55"
   emotion_data_v2$older[emotion_data_v2$age<30]<-"Under 30"
     emotion_data_v2|> 
  dplyr::rename("Enjoyment" = emotions_yesterday_1, "Physical_pain" = emotions_yesterday_2, "Worry" = emotions_yesterday_3, "Sadness" = emotions_yesterday_4 , "Stress" = emotions_yesterday_5,  "Anger" = emotions_yesterday_6, "Happiness" = emotions_yesterday_7)|>
  ggplot(mapping = aes(x = Week, y = Sadness, group = older, color = older))+ 
  stat_summary(geom="point", fun=mean)+
  geom_smooth()+ 
  geom_vline(xintercept = 20, linetype = "solid", color = "Purple", size = 0.5)+
  geom_vline(xintercept = 30, linetype = "solid", color = "Purple", size = 0.5)+
   geom_vline(xintercept = 64, linetype = "dotted", color = "red", size = 1)
       
```
### My observation
+ People became sadder when the lockdown started.
+ The drop around week 45 is when the COVID Vaccine was announced

## Happiness
```{r}
 emotion_data_v2|>
  dplyr::rename("Enjoyment" = emotions_yesterday_1, "Physical_pain" = emotions_yesterday_2, "Worry" = emotions_yesterday_3, "Sadness" = emotions_yesterday_4 , "Stress" = emotions_yesterday_5,  "Anger" = emotions_yesterday_6, "Happiness" = emotions_yesterday_7)|>
  ggplot(mapping = aes(x = Week, y = Happiness, group = older, color = older))+
  stat_summary(geom="point", fun=mean)+
  geom_smooth()+
  geom_vline(xintercept = 20, linetype = "solid", color = "Purple", size = 0.5)+
  geom_vline(xintercept = 30, linetype = "solid", color = "Purple", size = 0.5)+
   geom_vline(xintercept = 64, linetype = "dotted", color = "red", size = 1)
```
+ Happiness among people under 30 was pretty stable
+ When the lockdown started, the older 2 age groups became less happy

## Enjoyment
```{r}
emotion_data_v2|>
  dplyr::rename("Enjoyment" = emotions_yesterday_1, "Physical_pain" = emotions_yesterday_2, "Worry" = emotions_yesterday_3, "Sadness" = emotions_yesterday_4 , "Stress" = emotions_yesterday_5,  "Anger" = emotions_yesterday_6, "Happiness" = emotions_yesterday_7)|>
  ggplot(mapping = aes(x = Week, y = Happiness, group = older, color = older))+
  stat_summary(geom="point", fun=mean)+
  geom_smooth()+
  geom_vline(xintercept = 20, linetype = "solid", color = "Purple", size = 0.5)+
  geom_vline(xintercept = 30, linetype = "solid", color = "Purple", size = 0.5)+
   geom_vline(xintercept = 64, linetype = "dotted", color = "red", size = 1)
```

## Anger
```{r}
emotion_data_v2|>
  dplyr::rename("Enjoyment" = emotions_yesterday_1, "Physical_pain" = emotions_yesterday_2, "Worry" = emotions_yesterday_3, "Sadness" = emotions_yesterday_4 , "Stress" = emotions_yesterday_5,  "Anger" = emotions_yesterday_6, "Happiness" = emotions_yesterday_7)|>
  ggplot(mapping = aes(x = Week, y = Anger, group = older, color = older))+
  stat_summary(geom="point", fun=mean)+
  geom_smooth()+
  geom_vline(xintercept = 20, linetype = "solid", color = "Purple", size = 0.5)+
  geom_vline(xintercept = 30, linetype = "solid", color = "Purple", size = 0.5)+
   geom_vline(xintercept = 64, linetype = "dotted", color = "red", size = 1)
```
+ Younger people were angrier about the pandemic

## Stress
```{r}
emotion_data_v2|>
  dplyr::rename("Enjoyment" = emotions_yesterday_1, "Physical_pain" = emotions_yesterday_2, "Worry" = emotions_yesterday_3, "Sadness" = emotions_yesterday_4 , "Stress" = emotions_yesterday_5,  "Anger" = emotions_yesterday_6, "Happiness" = emotions_yesterday_7)|>
  ggplot(mapping = aes(x = Week, y = Stress, group = older, color = older))+
  stat_summary(geom="point", fun=mean)+
  geom_smooth()+
  geom_vline(xintercept = 20, linetype = "solid", color = "Purple", size = 0.5)+
  geom_vline(xintercept = 30, linetype = "solid", color = "Purple", size = 0.5)+
   geom_vline(xintercept = 64, linetype = "dotted", color = "red", size = 1)
```
+ Younger people were more stressed

## Worry
```{r}
emotion_data_v2|>
  dplyr::rename("Enjoyment" = emotions_yesterday_1, "Physical_pain" = emotions_yesterday_2, "Worry" = emotions_yesterday_3, "Sadness" = emotions_yesterday_4 , "Stress" = emotions_yesterday_5,  "Anger" = emotions_yesterday_6, "Happiness" = emotions_yesterday_7)|>
  ggplot(mapping = aes(x = Week, y = Worry, group = older, color = older))+
  stat_summary(geom="point", fun=mean)+
  geom_smooth()+
  geom_vline(xintercept = 20, linetype = "solid", color = "Purple", size = 0.5)+
  geom_vline(xintercept = 30, linetype = "solid", color = "Purple", size = 0.5)+
   geom_vline(xintercept = 64, linetype = "dotted", color = "red", size = 1)
```
+ People were the most worried during the lockdown

## Physical pain
```{r}
emotion_data_v2|>
  dplyr::rename("Enjoyment" = emotions_yesterday_1, "Physical_pain" = emotions_yesterday_2, "Worry" = emotions_yesterday_3, "Sadness" = emotions_yesterday_4 , "Stress" = emotions_yesterday_5,  "Anger" = emotions_yesterday_6, "Happiness" = emotions_yesterday_7)|>
  ggplot(mapping = aes(x = Week, y = Physical_pain, group = older, color = older))+
  stat_summary(geom="point", fun=mean)+
  geom_smooth()+ 
  geom_vline(xintercept = 20, linetype = "solid", color = "Purple", size = 0.5)+
  geom_vline(xintercept = 30, linetype = "solid", color = "Purple", size = 0.5)+
   geom_vline(xintercept = 64, linetype = "dotted", color = "red", size = 1)
```
+  Physical pain did not change much over time.

## Conclusions
+ COVID strongly affected people's emotions. 
+ People were more affected in the beginning of the pandemic than later when the death rate was higher.
+ Younger people's emotions were more affected.